---
title: "Data 624 - HW 1"
author: "Coco Donovan"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading Necessary Packages:

```{r}
library(fpp3)
library(forecast)
library(ggplot2)
```

### HA - 2.1

Explore the following four time series: Bricks from aus_production, Lynx from pelt, Close from gafa_stock, Demand from vic_elec.

#### a & b.

Use ? (or help()) to find out about the data in each series.
What is the time interval of each series?

##### aus_production:

```{r}
help("aus_production")

# aus_production
```

The time interval of aus_production is quarterly, meaning every three months.

##### pelt:

```{r}
help('pelt')

# pelt
```

The pelt data frame has a time interval of yearly observations.

##### gafa_stock:

```{r}
help('gafa_stock')

# gafa_stock

# unique(gafa_stock$Symbol)
```

The data frame "gafa_stock" contains observations on a the basis of a daily time 
interval.

##### vic_elec:

```{r}
help('vic_elec')

# vic_elec
```

The data frame "vic_elec" contains observations on a time interval of every half hour.

#### c.

Use autoplot() to produce a time plot of each series.

##### Bricks from aus_production:

```{r}
autoplot(aus_production, Bricks)
```

##### Lynx from pelt:

```{r}
autoplot(pelt, Lynx)
```

##### Close from gafa_stock:

```{r}
autoplot(gafa_stock, Close)
```

##### Demand from vic_elec:

```{r}
autoplot(vic_elec, Demand)
```

#### d.

For the last plot, modify the axis labels and title.

##### Demand from vic_elec:

```{r}
autoplot(vic_elec) + labs(title = "Electricity Demand in Victoria on a Half-Hourly Basis")+ ylab("Demand (MWh)")
```


